---
title: "S1.10 Code used to develop and present GLMM results"
author: "Clare Vernon"
date: "2025-03-11"
output: html_document
editor_options: 
  chunk_output_type: console
---

#### Download files and set working directory
The files are available on GitHub, here: https://github.com/ClareMVernon/ReviewShrubAlpineAustralia

Change the working directory as required and load the data file. This file contains data for both the 
```{r}
library(readr)
df_alldata <- read_csv("~/S2.4. Data of mean shrub cover-abundance used to construct GLMM.csv")
view(df_alldata)
```


#### Install packages and set working directory
```{r}
#oo <- options(repos = "https://cran.r-project.org/")
#utils::install.packages("Matrix")
#utils::install.packages("lme4")
#options(oo)
#utils::install.packages("https://cran.r-project.org/bin/windows/contrib/4.4/Matrix_1.7-0.zip", repos = NULL)
#utils::install.packages("https://cran.r-project.org/bin/windows/contrib/4.4/lme4_1.1-35.5.zip", repos = NULL)
#install.packages("mvabund")
#install.packages("MASS")
#install.packages("readxl")
#install.packages("ggplot2")
#install.packages("dplyr")
#install.packages("maps")
#install.packages("tidyr")
#install.packages("reshape2")
#install.packages("stringr")
#install.packages("ggrepel")
#install.packages("see")
#install.packages("devtools")
#devtools::install_github("thomasp85/patchwork")
#install.packages("PerformanceAnalytics")
#install.packages("faraway")
#install.packages("corrplot")
#install.packages(c("raster", "terra", "lubridate", "patchwork"))
```

Load other packages.
```{r}
library(lme4)
library(Matrix)
library(mvabund)
library(MASS)
library(readxl)
library(ggplot2)
library(dplyr)
library(maps)
library(tidyr)
library(reshape2)
library(stringr)
library(see)
library(patchwork)
library(PerformanceAnalytics)
library(raster)
library(terra)
library(lubridate)
library(patchwork)
library(faraway)
library(corrplot)
library(FSA) # required for Dunns test of significant differences
library(merDeriv)
library(parameters)
library(DHARMa)
```


#### 1. Climate GLMM Analysis ####
Note that the dataset used provides data for both the climate and fire-related GLMM. See 2. Fire GLMM Analysis below. 
```{r}
df_model  <- df_alldata %>%
  filter(!is.na(StandardisedAnnualChange )) # Remove NA values
df_modela <- df_model %>% filter (Climate == "Yes")# Extracts the fire-only related rows 
df_model <- df_modela %>% filter(Design == "BACI" | Design == "BA") # Omit CI designs, as this doesn't capture rate of change as t0 is not known. 

df_model$StandardisedAnnualChange <- as.numeric(df_model$StandardisedAnnualChange)
df_model$color <- with(df_model, ifelse(StandardisedAnnualChange > 0, "positive", ifelse(StandardisedAnnualChange < 0, "negative", "zero")))

# Ensure all data is correctly formatted. Note some values scaled for GLMM development. 
df_model$Year.Final <- as.numeric(df_model$Year.Final)
df_model$Year.Initial <- as.numeric(df_model$Year.Initial)
df_model$TimeSpan <- as.numeric(df_model$TimeSpan)
df_model$Elev_Est_Upper <- as.numeric(df_model$Elev_Est_Upper)
df_model$Elev_Est_Lower <- as.numeric(df_model$Elev_Est_Lower)
df_model$StandardisedAnnualChange.2<- as.numeric(df_model$StandardisedAnnualChange.2)
df_model$Elev_Est_Upper <- as.numeric(df_model$Elev_Est_Upper)
df_model$RainJan <- as.numeric(df_model$RainJan)
df_model$RainJuly<- as.numeric(df_model$RainJuly)
df_model$JanTempMax <- as.numeric(df_model$JanTempMax)
df_model$JanTempMin <- as.numeric(df_model$JanTempMin)
df_model$JulyTempMax <- as.numeric(df_model$JulyTempMax)
df_model$JulyTempMin <- as.numeric(df_model$JulyTempMin)
df_model$Elev_Est_Lowers <- scale(df_model$Elev_Est_Lower)
df_model$Elev_Est_Uppers <- scale(df_model$Elev_Est_Upper)
df_model$Year.Initial <- as.numeric(df_model$Year.Initial)
df_model$Year.Final <- as.numeric(df_model$Year.Final)
df_model$Year.Initials <- scale(df_model$Year.Initial)
df_model$Year.Finals  <- scale(df_model$Year.Final)
```

Testing whether underlying data follows a normal distribution. 
```{r}
qqnorm(df_model$StandardisedAnnualChange.2) # Non-linear, not a straight 45 degree line
shapiro.test(df_model$StandardisedAnnualChange.2) # p < 0.05, non-linear
df_model$Year.Final <- as.numeric(df_model$Year.Final)
m_linear <- glmer(StandardisedAnnualChange.2 ~ Year.Final + (1 | Response.Level), data = df_model)
```

Outliers are present. Test whether removing outliers improves normal distribution.
```{r}
# Outliers are presented. Test whether removing outliers improves normal distribution. 
df_model$StandardisedAnnualChange.2 <- as.numeric(df_model$StandardisedAnnualChange.2)
shapiro.test(df_model$StandardisedAnnualChange.2)

# Linear model of all data 
lm.test.normalresid <- lm(StandardisedAnnualChange.2 ~ Year.Final, data = df_model)
par(mfrow = c(2, 2))
plot(lm.test.normalresid)
par(mfrow = c(1, 1))
cooksd <- cooks.distance(lm.test.normalresid)
sample_size <- nrow(df_model)
plot(cooksd, pch = "*", cex = 2, main = "Influential Obs by Cook's Distance")
abline(h = 4/sample_size, col = "red")  # Add cutoff line
text(x = seq_along(cooksd), y = cooksd, labels = ifelse(cooksd > 4/sample_size, seq_along(cooksd), ""), col = "red")
influential <- which(cooksd > (4/sample_size))

# Create a new dataframe without outliers
df_model_no_outlier <- df_model[-influential, ] 
head(df_model_no_outlier)
hist(df_model_no_outlier$StandardisedAnnualChange.2)
hist(df_model$StandardisedAnnualChange.2)
lm.test.normalresid <- lm(StandardisedAnnualChange.2 ~ Year.Final, data = df_model_no_outlier)
par(mfrow = c(2, 2))
plot(lm.test.normalresid ) # Look at the results
par(mfrow = c(1,1))

# Does the normality improve if outliers are removed? 
shapiro.test(df_model$StandardisedAnnualChange.2) # p < 0.05
shapiro.test(df_model_no_outlier$StandardisedAnnualChange) # p < 0.05
```
Outcome: No significant improvement, p << 0.001 for both models. Retain all datasets to better represent fire responses of shrubs.

Assess correlations. 
Highly correlated variables (r > 0.70) are to be excluded from the same model as model terms. 
```{r}
subset_corrdata <- df_model[, c("StandardisedAnnualChange.2",
                                "Elev_Est_Lower", 
                                "Elev_Est_Upper", 
                                "Year.Initial", 
                                "Year.Final",
                                "RainJan", 
                                "RainJuly", 
                                "JanTempMax", 
                                "JanTempMin", 
                                "JulyTempMax", 
                                "JulyTempMin")]
res <- cor(subset_corrdata, method = c("spearman"))
corrplot(res, tl.col = "red", bg = "White", tl.srt = 35, 
         title = "\n\n Correlation plot of continous variables \n",
         addCoef.col = "black", type = "lower") # Show correlation strengths. Same plot shown in S3 with relabelled variables for clarity. 
p_corrchart <- chart.Correlation(subset_corrdata[,1:11], histogram=FALSE, method = c("spearman"), pch="19", exact = FALSE)
p_corrchart # Alternative correlation plot showing the significance and distribution of data. 
```

Assessing whether data differs based on study characteristics. 
Using Kruskal-Wallis test as the assumption of normality is violated, and Dunn's test for pairwise comparisions if a significant difference is detected. 
```{r}
# Population Scale
kruskal.test(StandardisedAnnualChange.2 ~ Response.Level, data = df_model)

# Alpine zone of shrub species
kruskal.test(StandardisedAnnualChange.2 ~ Alpine.Zone, data = df_model)

# Ecosystem type 
kruskal.test(StandardisedAnnualChange.2 ~ Aggregate.InitialES, data = df_model)
dunnTest(StandardisedAnnualChange.2 ~ Aggregate.InitialES, data = df_model, method="bh")
```

Testing model requirements. 
Test requirement 1. That > 10-20 samples per treatment -  this is met as all samples have temperature, rainfall and final sampling year. 
Test requirement 2. That categorical variables have at least six levels.
```{r}
n_distinct(df_model$Study) # > 6
n_distinct(df_model$Studysite) # > 6
n_distinct(df_model$Response.Level) # < 6 levels, omit
```

#### 1.1. Development of Full Climate Model #### 
```{r}
# Null model 
m.null <- glmer(StandardisedAnnualChange.2 ~ (1 | Study) + (1 | Studysite) + (1 | Year.Finals) + (1 | Elev_Est_Lower), data = df_model)
summary(m.null)
AIC(m.null)

# Overfitted model 
m.overfit <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals) + (1 | Study) + (1 | Studysite) + (1 | Elev_Est_Lower), data = df_model)
summary(m.overfit) # REML = 357.2
AIC(m.overfit)

# Test REML of all possible model combinations 
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals) + (1 | Study) + (1 | Studysite) + (1 | Elev_Est_Lower), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Study), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Studysite), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Elev_Est_Lower), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals) + (1 | Study), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals) + (1 | Studysite), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals) + (1 | Elev_Est_Lower), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Study) + (1 | Studysite), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Studysite) + (1 | Elev_Est_Lower), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Study) + (1 | Elev_Est_Lower), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals)+ (1 | Studysite) + (1 | Elev_Est_Lower), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals) + (1 | Study) + (1 | Studysite), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals) + (1 | Study) + (1 | Elev_Est_Lower), data = df_model)
summary(m.x)
m.x <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Study) + (1 | Studysite) + (1 | Elev_Est_Lower), data = df_model)
summary(m.x)

# Reduce by removing RE with variance = 0
# Elev_Est_Lower first
m.reduce1 <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals) + (1 | Study) + (1 | Studysite), data = df_model)
summary(m.reduce1)
AIC(m.reduce1) # AIC improved 

# Reduce further, test AIC and variance when studysite is removed
m.reduce2 <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals) + (1 | Study), data = df_model)
summary(m.reduce2)
AIC(m.reduce2) # AIC note improvement, AIC > 10

# Reduce further, test AIC and variance when studysite is retained, study ID removed
m.reduce3 <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals) + (1 | Studysite), data = df_model) # FINAL MODEL CHOSEN
summary(m.reduce3)
AIC(m.reduce3) # AIC note improvement, AIC within 2 units of m.reduce1. This will be the chosen model without interaction effect. 

# Reduce further, test AIC and variance when studysite is retained, study ID retained, remove year
m.reduce4 <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Study) + (1 | Studysite), data = df_model)
summary(m.reduce4)
AIC(m.reduce4) # AIC note improvement, AIC within 2 units of m.reduce1

# Reduce further, test AIC and variance when studysite is retained, remove study ID retained and remove year
m.reduce5 <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Studysite), data = df_model)
summary(m.reduce5)
AIC(m.reduce5) # AIC note improvement, AIC within 2 units of m.reduce1

# Assessment of non-interaction model performance
m.final.noninteract <- m.reduce3
model_parameters(m.final.noninteract )
AIC(m.final.noninteract)
anova(m.final.noninteract )
model_parameters(m.final.noninteract )

### Creating and assessing interaction terms
m.reduce1.interaction.0 <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin + RainJan + RainJuly + (1 | Year.Finals) + (1 | Studysite), data = df_model) 
m.reduce1.interaction.1 <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin * RainJuly + RainJan  + (1 | Year.Finals) + (1 | Studysite), data = df_model)  # Model 8
m.reduce1.interaction.2 <- glmer(StandardisedAnnualChange.2 ~ JanTempMax * RainJan + JulyTempMin  + RainJuly + (1 | Year.Finals) + (1 | Studysite), data = df_model)

AIC(m.reduce1.interaction.0, m.reduce1.interaction.1, m.reduce1.interaction.2)
model_parameters(m.reduce1.interaction.0)
model_parameters(m.reduce1.interaction.1)
model_parameters(m.reduce1.interaction.2) # JulyTempMin significant, as is interaction effect chosen model 

# Final model is m.reduce1.interaction.1 <- glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin * RainJuly + RainJan  + (1 | Year.Finals) + (1 | Studysite), data = df_model)
m.final <-  glmer(StandardisedAnnualChange.2 ~ JanTempMax + JulyTempMin * RainJuly + RainJan  + (1 | Year.Finals) + (1 | Studysite), data = df_model)
summary(m.final)
model_parameters(m.final)
AIC(m.final, m.reduce1.interaction.0)
anova(m.final, m.reduce1.interaction.0)

# Testing performance of final model 
MuMIn::r.squaredGLMM(m.final)
sim_res <- simulateResiduals(m.final)
plot(sim_res)
testUniformity(sim_res)
testDispersion(sim_res) # No evidence of problematic dispersion, use Wald Z and chi-sq tests
model_parameters(m.final)
summary(m.final)
AIC(m.final)
```

#### 1.2. Development of Reduced Climate Model. ####
```{r}
m.final2 <-  glmer(StandardisedAnnualChange.2 ~ JulyTempMin * RainJuly + (1 | Year.Finals) + (1 | Studysite), data = df_model)
model_parameters(m.final2) 
summary(m.final2)
AIC(m.final, m.final2)

# Testing performance of final model 
MuMIn::r.squaredGLMM(m.final2)
sim_res <- simulateResiduals(m.final2)
plot(sim_res)
testUniformity(sim_res)
testDispersion(sim_res) # No evidence of problematic dispersion, use Wald Z and chi-sq tests
model_parameters(m.final2)
summary(m.final2)

AIC(m.final) 
AIC(m.final2)
AIC(m.final) - AIC( m.final2)

# Model validation: check that AIC increases when fixed effect terms are removed
m.final2.NoInteract <-  glmer(StandardisedAnnualChange.2 ~ JulyTempMin + RainJuly + (1 | Year.Finals) + (1 | Studysite), data = df_model)
AIC(m.final2.NoInteract )
model_parameters(m.final2.NoInteract )

m.final2.reduce.RainJuly <-  glmer(StandardisedAnnualChange.2 ~ RainJuly + (1 | Year.Finals) + (1 | Studysite), data = df_model)
AIC(m.final2.reduce.RainJuly )
model_parameters(m.final2.reduce.RainJuly )

m.final2.JulyTemp <-  lmer(StandardisedAnnualChange.2 ~ JulyTempMin  + (1 | Year.Finals) + (1 | Studysite), data = df_model)
AIC(m.final2.JulyTemp)
model_parameters(m.final2.JulyTemp )

AIC(m.final2)
model_parameters(m.final2)

```

Plotting the results
```{r}
median(df_model$JulyTempMin)
p1.JulyTemp <- ggplot (data = df_model, aes(y = StandardisedAnnualChange.2, x = JulyTempMin)) + geom_point() + geom_smooth(method = lm) + theme_classic() + xlab("Average monthly minimum temperature °C, July (winter)") + ylab("Annual change cover-abundance (%)") + ggtitle("a") + geom_vline(xintercept = -1.5, colour="grey", linetype = "longdash", size= 1)  +  geom_hline(yintercept = 0, colour="red", linetype = "longdash", size= 1)
p1.JulyTemp

median(df_model$RainJuly)
p1.JulyRain <- ggplot (data = df_model, aes(y = StandardisedAnnualChange.2, x = RainJuly)) + geom_point() + geom_smooth(method = lm) + theme_classic() + xlab("Total monthly precipitation, mm, July (winter)") + ylab("Annual change cover-abundance (%)") + ggtitle("b") + geom_vline(xintercept = 118.2904, colour="grey", linetype = "longdash", size= 1)  +  geom_hline(yintercept = 0, colour="red", linetype = "longdash", size= 1)
p1.JulyRain

p1.cover <- ggplot (data = df_model, aes(y = StandardisedAnnualChange.2, x = as.numeric(Year.Final))) + geom_point() + geom_smooth(method = lm) + theme_classic() + xlab("Year") + ylab("Annual change cover-abundance (%)") +  geom_hline(yintercept = 0, colour="red", linetype = "longdash", size= 1) + ggtitle("c")
p1.cover

p1.JulyTemp | p1.JulyRain | p1.cover

# Cover change through time 
corr <- cor.test(x=df_model$Year.Final, y=df_model$StandardisedAnnualChange.2, method = 'spearman')
corr
```


#### 2. Fire GLMM Analysis ####
Select relevant studies from the dataset related to fire.
```{r}
df_model <- df_alldata %>% filter (Treatment_Thematic == "Fire" |
                      Treatment_Thematic == "Fire + Grazing" | 
                       Treatment_Thematic ==  "Fire + Restoration" |
                      Treatment_Thematic == "Fire + Climate change" ) # Extracts the fire-only related rows 
df_model <- df_model %>% filter(Design == "BACI" | Design == "BA") # Omit CI designs, as this doesn't capture rate of change as t0 is not known. 
df_model <- df_model %>% filter(Control.Exposure == "Exposure") # Omit Control i.e. unburnt sites. 
df_model$TimeSinceFire <- df_model$TimeSinceFire.New # Rename to something more sensible 
```

Ensure that all variables are interpreted as the correct data type. 
```{r}
df_model$TimeSinceFire <- as.numeric(df_model$TimeSinceFire)
df_model$StandardisedAnnualChange <- as.numeric(df_model$StandardisedAnnualChange)
df_model$Year.Final <- as.numeric(df_model$Year.Final)
df_model$Year.Initial <- as.numeric(df_model$Year.Initial)
df_model$TimeSpan <- as.numeric(df_model$TimeSpan)
df_model$TimeSinceFire <- as.numeric(df_model$TimeSinceFire)
df_model$Elev_Est_Upper <- as.numeric(df_model$Elev_Est_Upper)
df_model$Elev_Est_Lower <- as.numeric(df_model$Elev_Est_Lower)
df_model$TimeSinceFire <- as.numeric(df_model$TimeSinceFire)
df_model$StandardisedAnnualChange.2<- as.numeric(df_model$StandardisedAnnualChange.2)
```

Identify outliers, and test whether removing outliers improves assumption of normality.
```{r}
# All values included
lm.test.normalresid <- lm(StandardisedAnnualChang2e ~ TimeSinceFire, data = df_model)
par(mfrow = c(2, 2))
plot(lm.test.normalresid)
par(mfrow = c(1,1))
cooksd <- cooks.distance(lm.test.normalresid)
sample_size <- nrow(df_model)
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4/sample_size, col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4/sample_size, names(cooksd),""), col="red")  # add labels

# Identify and remove outliers in a new dataframe
influential <- as.numeric(names(cooksd)[(cooksd > (4/sample_size))])
df_model_no_outlier <- df_model[-influential, ] # Test dataframe
lm.test.normalresid.no.outlier <- lm(StandardisedAnnualChange ~ TimeSinceFire, data = df_model_no_outlier) # Rerun model 
par(mfrow = c(2, 2))
plot(lm.test.normalresid.no.outlier) # Look at the results
par(mfrow = c(1,1))

shapiro.test(df_model$StandardisedAnnualChange)
shapiro.test(df_model_no_outlier$StandardisedAnnualChange)
```
Outcome: both dataframes are non-normal (p << 0.001). Decision to retain all data for full representation of results. 

Test of normal distribution of residuals.
```{r}
glm.test.normalresid <- glm(StandardisedAnnualChange.2~ TimeSinceFire, data = df_model)
par(mfrow = c(2, 2))
plot(lm.test.normalresid)
par(mfrow = c(1,1))
```

Assessing whether data differs based on study characteristics. 
Using Kruskal-Wallis test as the assumption of normality is violated, and Dunn's test for pairwise comparisions if a significant difference is detected. 
```{r}
# Alpine zone of shrub species
kruskal.test(StandardisedAnnualChange.2 ~ Alpine.Zone, data = df_model)

# Ecosystem type 
kruskal.test(StandardisedAnnualChange.2 ~ Aggregate.InitialES, data = df_model)
dunnTest(StandardisedAnnualChange.2 ~ Aggregate.InitialES, data = df_model, method="bh")
```

Testing model requirements. 
Test requirement 1. That > 10-20 samples per treatment -  this is met as all samples have temperature, rainfall and final sampling year. 
Test requirement 2. That categorical variables have at least six levels.
```{r}
n_distinct(df_model$Study) # > 6
n_distinct(df_model$Studysite) # > 6
n_distinct(df_model$Response.Level) # < 6 levels, omit
```

Correlation plot of numerical data. 
```{r}
subset_corrdata <- df_model[, c("StandardisedAnnualChange.2",
                                "Elev_Est_Lower", 
                                "Elev_Est_Upper", 
                                "Year.Initial", 
                                "Year.Final", 
                                "TimeSinceFire",
                                "RainJan", 
                                "RainJuly", 
                                "JanTempMax", 
                                "JanTempMin", 
                                "JulyTempMax", 
                                "JulyTempMin")]
res <- cor(subset_corrdata)
round(res, 2)
p_corrchart <- chart.Correlation(subset_corrdata[,1:12], histogram=TRUE, method = c("spearman"), pch="19", exact = FALSE)
```

#### 2.1. Fire Model
```{r}
m1.null <- lmer(StandardisedAnnualChange ~ 1 + (1 | JanTempMax) + (1 | Year.Final) + (1 | RainJuly) + (1 | Elev_Est_Lower), data = df_model, REML = FALSE)
summary(m1.null)
anova(m1.null)
m.final <- m1.null

# Testing performance of final model 
MuMIn::r.squaredGLMM(m.final)
sim_res <- simulateResiduals(m.final)
plot(sim_res)
testUniformity(sim_res)
testDispersion(sim_res) # No evidence of problematic dispersion, use Wald Z and chi-sq tests
model_parameters(m.final)
summary(m.final)

```

#### 2.2. Fire Trait Model
```{r}
# Forward selection procedure for random effects
m1 <- glmer(StandardisedAnnualChange ~ TimeSinceFire + Seeder.Resprouter2 + (1 | JanTempMax) + 1, data = df_model)
m2 <- glmer(StandardisedAnnualChange ~ TimeSinceFire + Seeder.Resprouter2 + (1 | JanTempMax) + (1 | Year.Final) + 1, data = df_model)
m3 <- glmer(StandardisedAnnualChange ~ TimeSinceFire + Seeder.Resprouter2 + (1 | JanTempMax) + (1 | Year.Final) + (1 | RainJuly) + 1, data = df_model)
m4 <- glmer(StandardisedAnnualChange ~ TimeSinceFire + Seeder.Resprouter2 + (1 | JanTempMax) + (1 | RainJuly) + 1, data = df_model)
m5 <- glmer(StandardisedAnnualChange ~ TimeSinceFire + Seeder.Resprouter2 + (1 | JanTempMax) + (1 | Year.Final) + (1 | RainJuly) + (1 | Elev_Est_Lower) + 1, data = df_model)
m6 <- glmer(StandardisedAnnualChange ~ TimeSinceFire + Seeder.Resprouter2 + (1 | Year.Final) + (1 | RainJuly) + (1 | Elev_Est_Lower) + 1, data = df_model)
m7 <- glmer(StandardisedAnnualChange ~ TimeSinceFire + Seeder.Resprouter2 + (1 | JanTempMax) + (1 | Year.Final) + (1 | Elev_Est_Lower) + 1, data = df_model)
m8 <- glmer(StandardisedAnnualChange ~ TimeSinceFire + Seeder.Resprouter2 + (1 | Year.Final) + (1 | RainJuly) + (1 | Elev_Est_Lower) + 1, data = df_model)
m9 <- glmer(StandardisedAnnualChange ~ TimeSinceFire + Seeder.Resprouter2 + (1 | RainJuly) + (1 | Elev_Est_Lower) + 1, data = df_model)

AIC(m1, m2, m3, m4, m5, m6, m7, m8, m9, m1.null) # m.null lowest AIC, with m9 lowest AIC of all Fire Trait Model iterations.
anova(m1, m2, m3, m4, m5, m6, m7, m8, m9, m1.null)

summary(m9)
m.final <- m9
```

```{r}
# Testing performance of final model 
MuMIn::r.squaredGLMM(m.final)
sim_res <- simulateResiduals(m.final)
plot(sim_res)
testUniformity(sim_res)
testDispersion(sim_res) # No evidence of problematic dispersion, use Wald Z and chi-sq tests
model_parameters(m.final)
summary(m.final)
```

